---
title: 'Exercise 2: Data types, structures and conditions'
author: "Sine Zambach"
date: "September 12, 2024"
output: 
  html_document:
    theme: sandstone
    code_folding: show
editor_options: 
  markdown: 
    wrap: 72
---

This is a RMD-file (An R-MarkDown file) that you can use for your report
and exercises. You should start by opening the Rmd-file in RStudio. If
you then click the *Knit* button, you will get the HTML result shown in
the Viewer pane in RStudio.

HINT: You can find code to help you in the compendium or in the slides.

# 1. Vectors

## Creation and initialization

-   Use c() to initialize different vectors of length 3 but with
    different data types; numeric (double), numeric (integer), logical
    and character.

```{r}

v1 <- c(1,2,3)
v2 <- c(T,F,T)
v3 <- c(10L,20L,30L)
v4 <- c("hello", "hi", "bye")
v5 <- c(1.3, 2.4, 3.5)

class(v1)
class(v2)
class(v3)
class(v5)

typeof(v3)

```

## Inspection

-   Examine the vectors by using class(), length() and str(), typeof().
    Do it in the code chunk above or create a new one yourself

## Manipulation

We retrieve values in a vector by declaring an index inside a single
square bracket "[]" operator. In other words we tell R to give us the
i'th element in the vector.

-   For the below "sorted_vector", get the first value (i=1)

-   Replace the second element with e.g. 100

```{r}
sorted_vector <- seq(from = 1.5, to = 5, by = 0.5)
length(sorted_vector)

# first element?
sorted_vector[1]

# replace second element with 100
sorted_vector[2] <- 100
sorted_vector
```

Append a value to one of your vectors

```{r}

# Insert code here
sorted_vector <- append(sorted_vector,6, after = 0)


```

-   Add two numeric vectors together. Inspect the result.
-   Multiply two numeric vectors together. Inspect the results

```{r}
# Insert code 
sorted_vector + v1
v1 + v3

v1 * v3

```

# 2. Data Frames

In data frames each row corresponds to e.g. an individual and each
column corresponds to a variable. Just as for lists, you can store
different data types in data frames.

## Create

Create a data frame with the below objects as columns using
data.frame().

```{r}
temp <- c(20.37, 18.56, 18.4, 21.96, 29.53, 28.16,
36.38, 36.62, 40.03, 27.59, 22.15, 19.85)
humidity <- c(88, 86, 81, 79, 80, 78, 71, 69, 78, 82, 85, 83)
rain <- c(72, 33.9, 37.5, 36.6, 31.0, 16.6,
1.2, 6.8, 36.8, 30.8, 38.5, 22.7)
month <- c("January", "February", "March", "April", "May", "June",
"July", "August", "September", "October", "November", "December")

# write your code here
DF <- data.frame(temp, humidity, rain, month)


```

## Inspect and manipulate

Inspect your data frame by using
`nrow(), ncol(), dim(), head(), tail(), str(), names()`

For your data frame use the "`$[column_name]`" to access a column

For your data frame use the "`[,i]`" to access a column

Advanced: Add a new column to your data frame by using the syntax:
`[dataframe]$[new_column_name] <- [new_column]`

```{r}

#Inspect
nrow(DF)
ncol(DF)
dim(DF)

head(DF, 10)
tail(DF)

str(DF)
names(DF)
summary(DF)
View(DF)

# access a column:
DF[,1] # Like the A-column in Excel
DF$temp # same - using another notation
DF[,"temp"] # Like the A-column in Excel

# Acessing rows
DF[1,] # Like first row in Excel
DF[1,3] # Like C1 in Excel. 



```

## Subsetting data.frame

You can subset (you will learn another function for this later) where
the column X1.10 has to be above 4.

```{r}
# Create D
letters <- sample(LETTERS[1:3], 10, replace = TRUE)
D <- data.frame(1, 1:10, letters)


# Create subset SubD 
SubD <- D[D$X1.10 >= 5, ] # New DF with only C in "letter"



```

# 3. Conditions (if/else-statements)

Conditions makes a computer program operational. Therefore, we shall of
cause play around with it.

## Simple condition

-   Use the slides and write a simple condition that compares the two
    atomic vectors A and B. (prints "Yes" if equal, and "No" if not).

-   Try to change with letters in the alphabet. What happens?

```{r}
A <- "Cz"
B <- "Dd"
# write code here

if (A < B) {
  print("Yes")
} else {
   print("No")
}

ifelse(A < B,"Yes", "No")
ifelse(A == B,"Yes", "No")
```

-   Try to make an if-else condition that tests if `Atom` is in the
    vector `C` and prints "Yes" if so and "No" if not.

```{r}
Atom <- "YYou"   # change the values in Atom
C <- c("hello", "you", "me", "world")

# use %in%
# white code here
if (Atom %in% C) {
  print("Yes")
} else {
   print("No")
}

ifelse(Atom %in% C, "Yes", "No")

############
#  Check if both length of C is above 1 and if Atom is contained in C
# use &&
# use length()

if (length(C)>1 && Atom %in% C) {
  print("Yes")
} else {
   print("No")
}

ifelse(length(C)>1 && Atom %in% C, "Yes", "No")

```

-   Try to make a statements in which you both test if the length of `C`
    is above 1 and if `Atom` is contained in `C`. Use (&&) or (\|\|).

# 4. Groups for mandatory assignment

-   Find/create the last groups
-   Find data if you have not already

# 5. Extra exercises

## If you are happy

Here is a simple code that test if you are happy and then prints "clap
your hands" if so.

Now try to: - Change the value to other values such as FALSE, 2, 0, -1
and "yes". What happens? - Expand with an `else{}`-statement in which
you write something suitable such as "Have a hug", or similar. - Expand
with one or more `else{if{}}` and mood-statements. - Make two conditions
connected with and (&&) or or (\|\|). E.g. "you_know_it" should be
variable with a boolean value.

```{r}
happy <- TRUE

if(happy) {
  "clap your hands"
}

```

```{r}
happy <- F
angry <- F

if(happy) {
  "clap your hands"
}else{"stamp your feet"}

if(happy) {
  "clap your hands"
} else{if(angry)
  {"stamp your feet"}
}
```

## `if()` and `ifelse()`

What's the difference between `if` and `ifelse()`? Carefully read the
help and construct one examples that illustrate the key differences.

```{r}

```

## Subsetting data.frame - extra

BUILDS on: You can subset (you will learn another function for this
later) where the column X1.10 has to be above 4.

**Now** try to subset again, from `D`, where `letter` has to be `"B"`
and `X1.10` has to be *more than or equal* \`5

```{r}
# Create D
letters <- sample(LETTERS[1:3], 10, replace = TRUE)
D <- data.frame(1, 1:10, letters)

# Create subset ND

```

\`.
