---
title: "BA-BINTV2006U  Programming and Data Analysis for Business: Exercise 1"
author: "by Felix Thomsen, Alexander Schiøtz, Anders Laurents, Bertram Hage"
date: "2024-10-01"
output: 
  html_document:
    theme: sandstone
    code_folding: show
---

## 1. packages + dataload
<!-- 2. Load the dataset into R and inspect its structure. How many rows, how many columns, what is the
summary statistics for the numeric variables? -->
```{r init, warning = FALSE, message=FALSE}
library(knitr)
library(markdown)
library(rmarkdown)
library(tidyverse)
library(lubridate)
library(readxl)
theme_set(theme_bw())
```
```{r, message=FALSE}
data <- read_csv("data.csv")

# Remove outliers based on the IQR method for numeric columns
remove_outliers <- function(df, cols) {
  for (col in cols) {
    Q1 <- quantile(df[[col]], 0.25)
    Q3 <- quantile(df[[col]], 0.75)
    IQR <- Q3 - Q1
    df <- df %>% filter(df[[col]] >= (Q1 - 1.5 * IQR) & df[[col]] <= (Q3 + 1.5 * IQR))
  }
  return(df)
}

numeric_cols <- sapply(data, is.numeric)
data <- remove_outliers(data, names(data)[numeric_cols])
```

## Create a point plot or jitter plot
```{r}
#Jitter plot for hours studied and exam score
ggplot(data, aes(x = Hours_Studied, y = Exam_Score, color = Gender)) + geom_jitter() + scale_fill_manual(values = c("blue", "pink")) + labs(title = "Jitter plots of hours studied and exam scores + colored by Gender")
```

## Create a boxplot or a bar plot. 
To visualize the distribution of "Hours_Studied" by "Gender" we selected a boxplot:
```{r}
# data$Hours_Studied

ggplot(data, aes(x = data$Gender, y = data$Hours_Studied)) + geom_boxplot() +   labs(title = "Boxplot of Hours Studied by Gender", x = "Gender", y = "Hours Studied")
```

Generally we see that males and females studies the same amount of hours.

## Generate a histogram to visualize the distribution of one of the numerical variables.
```{r}

ggplot(data, aes(x = Exam_Score)) + geom_histogram(binwidth = 1, aes(fill = ..count..)) + scale_fill_gradient(low = "green", high = "blue") + labs(title = "Histogram of Exam Scores by Student", x = "Exam Score" , y = "Counts of Students")

```

## Create a visualization of own choice that you have “polished”,  and write what the purpose is with  your visualization. (Is it for a managerial decision? A publication? How?)
```{r}
ggplot(data, aes(x=Parental_Involvement, fill = Parental_Involvement))+
  geom_bar() +
  labs(title="Number of students sorted by Parental involvement", x="Parental involvement", y="Students")+
  ylim(0,4000) +
  theme(legend.position = "none",axis.text = element_text(size = 12))
```

## Create a linear model in which you model using at least tree input/independant variables. 
```{r}
print("anders er gay")
```
## Visualize the model output using dotwhiskers and stargazer
```{r}
print("anders er gay")
```

## Reflect on the model – how could it be improved? 

## Reflect on how you have used Generative AI (if you have).