---
title: "BA-BINTV2006U  Programming and Data Analysis for Business: Exercise 1"
author: "by Felix Thomsen, Alexander Schiøtz, Anders Laurents, Bertram Hage"
date: "2024-10-01"
output: 
  html_document:
    theme: sandstone
    code_folding: show
---

## 1. packages + dataload
<!-- 2. Load the dataset into R and inspect its structure. How many rows, how many columns, what is the
summary statistics for the numeric variables? -->
```{r init, warning = FALSE, message=FALSE}
library(knitr)
library(markdown)
library(rmarkdown)
library(tidyverse)
library(lubridate)
library(readxl)
theme_set(theme_bw())
```
```{r, message=FALSE}
data <- read_csv("data.csv")
```

## Create a point plot or jitter plot
```{r}
print("anders er gay")
```

## Create a boxplot or a bar plot. 
```{r}
print("anders er gay")
```
## Generate a histogram to visualize the distribution of one of the numerical variables.
```{r}
print("anders er gay")
```

## Create a visualization of own choice that you have “polished”,  and write what the purpose is with  your visualization. (Is it for a managerial decision? A publication? How?)
```{r}
print("anders er gay")
```

## Create a linear model in which you model using at least tree input/independant variables. 
```{r}
print("anders er gay")
```
## Visualize the model output using dotwhiskers and stargazer
```{r}
print("anders er gay")
```

## Reflect on the model – how could it be improved? 

## Reflect on how you have used Generative AI (if you have).